Scattering from compact objects: \\ Regge poles and the Complex Angular Momentum method

### 2019-12-17 rainbowCAM.tex
Tom, please add the new versions of the plots in Fig 1 to the repository.
made some comments on poles in the other quadrants

Next, we examine the role of Regge poles in the time-independent scattering of monochromatic planar waves. We apply Complex Angular Momentum techniques to re-sum the partial wave series for the scattering amplitude, expressing it as a residue series evaluated at poles in the first and fourth quadrants, accompanied by a background integral. We compute the scattering cross section at several frequencies, and show precise agreement with the partial-wave calculations. Finally, we show that compact bodies naturally give rise to orbiting, glory, and rainbow-scattering interference effects.
\end{abstract}

@@

 \begin{figure}[htb]
\centering
 \includegraphics[scale=0.50]{potentials_s02_CD_R6L2_v4} \\ \vspace{0.2cm}
 \includegraphics[scale=0.50]{potentials_s02_CD_R2p26L2_v4}
\caption{\label{RP_approx_2Mw_3_6_s_1} The effective potential $V_\ell$ for a quadrupole ($\ell =2$) perturbation of a compact body of constant density and tenuity $R/M = 6$ (upper) and $R /M = 2.26$ (lower). The scalar field potential - Eq.~(\ref{Inside_Potentiel}) - and axial gravitational-wave potential - Eq.~(\ref{RW_Potentiel_axialGW}) - are indicated as solid/black lines and dotted/red lines, respectively. The horizontal axis is the tortoise coordinate $r_\ast/(2M)$ defined in Eq.~(\ref{eq:tortoise}).}
\label{fig:Veff}
\end{figure}

@@

Figure \ref{fig:rp1} shows the Regge pole spectrum for a neutron-star-like body ($R/M = 6$). We see that there are two branches of Regge poles in the first quadrant, which meet at $\text{Re}( \lambda_n(\omega) ) \sim \omega b_0$, where $b_0$ is the impact parameter for the null ray which grazes the surface of the compact body. The axial GW $s=2$ modes (filled markers) are typically close to their scalar-field $s=0$ counterparts (unfilled markers), as might be anticipated from the similarities in the effective potentials for the two cases (see Fig.~\ref{fig:Veff}). \tom{We find no evidence of Regge poles in the fourth quadrant when $\omega>0$. For $\omega<0$ we find that $\lambda_n(-\omega) = \lambda_n(\omega)$. The contour integral deformation means we do not consider poles in the second and third quadrants (for Mohamed: I propose we do not go into more detail about this and the choice of boundary condition...but I am happy to be over ruled!)}

@@

Now, we ``deform'' the contour ${\cal C}$ in Eq.~(\ref{SW_Scalar_Scattering_amp}) in order to collect, by using Cauchy's theorem, the Regge poles contributions. This is achieved by following, \textit{mutatis mutandis}, the approach developed in Ref~\cite{Folacci:2019cmc} (see more particularly Sec. IIB~3 and Fig.~1). We obtain
%
\begin{equation}\label{CAM_Scalar_Scattering_amp_tot}
\hat{f} (\omega, \theta) =  \hat{f}^\text{\tiny{B}} (\omega, \theta) +  \hat{f}^\text{\tiny{RP}} (\omega, \theta)
\end{equation}
where
\begin{subequations}\label{CAM_Scalar_Scattering_amp_decomp}
\begin{equation}\label{CAM_Scalar_Scattering_amp_decomp_Background}
\hat{f}^\text{\tiny{B}} (\omega, \theta) = \hat{f}^\text{\tiny{B},\tiny{Re}} (\omega, \theta)+ \hat{f}^\text{\tiny{B},\tiny{Im}} (\omega, \theta)
\end{equation}
is a background integral contribution with
\begin{equation}\label{CAM_Scalar_Scattering_amp_decomp_Background_a}
\hat{f}^\text{\tiny{B},\tiny{Re}} (\omega, \theta) = \frac{1}{\pi \omega} \int_{{\cal C}_{-}} d\lambda \, \lambda S_{\lambda -1/2}(\omega) Q_{\lambda -1/2}(\cos \theta +i0)
\end{equation}
and
\begin{eqnarray}\label{CAM_Scalar_Scattering_amp_decomp_Background_b}
\hat{f}^\text{\tiny{B},\tiny{Im}} && (\omega, \theta) = \frac{1}{2 \omega}\left(\int_{+i\infty}^{0} d\lambda \, \left[S_{\lambda -1/2}(\omega) P_{\lambda-1/2} (-\cos \theta) \right. \right.\nonumber \\
&& -\left. \left. S_{-\lambda -1/2}(\omega) e^{i \pi \left(\lambda+1/2\right)}P_{\lambda-1/2} (\cos \theta) \right]\lambda\right).
\end{eqnarray}
\end{subequations}
The second term in Eq.~(\ref{CAM_Scalar_Scattering_amp_tot}) ,
\begin{eqnarray}\label{CAM_Scalar_Scattering_amp_decomp_RP}
& & \hat{f}^\text{\tiny{RP}} (\omega, \theta) = -\frac{i \pi}{\omega}    \sum_{n=1}^{+\infty}   \frac{ \lambda_n(\omega) r_n(\omega)}{\cos[\pi \lambda_n(\omega)]}  \nonumber \\
&&  \qquad\qquad \qquad\qquad \times  P_{\lambda_n(\omega) -1/2} (-\cos \theta),
\end{eqnarray}
is a sum over the Regge poles lying in the first and fourth quadrants of the CAM plane. Of course, Eqs.~(\ref{CAM_Scalar_Scattering_amp_tot}), (\ref{CAM_Scalar_Scattering_amp_decomp}) and (\ref{CAM_Scalar_Scattering_amp_decomp_RP}) provide an exact representation of the scattering amplitude $\hat{f} (\omega, \theta)$ for the scalar field, equivalent to the initial partial wave expansion (\ref{Scalar_Scattering_amp}). From this CAM representation, we can extract the contribution $\hat{f}^\text{\tiny{RP}} (\omega, \theta)$ given by (\ref{CAM_Scalar_Scattering_amp_decomp_RP}) which, as a sum over Regge poles, is only an approximation of $\hat{f} (\omega, \theta)$, and which provides us with a corresponding approximation of the differential scattering cross section via (\ref{Scalar_Scattering_diff}).

@@

We can be confident that there are no additional branches of Regge poles in the first or fourth quadrants, for two reasons. First, we have scanned across the complex-$\lambda$ domains shown in Figs.~\ref{RP_R6_approx_2Mw_3_6_s_1} and \ref{RP_R226_approx_2Mw_3_6_s_1}, as detailed in Sec.~\ref{subsec:method}. Second, in our calculation of $d\sigma/d\Omega$, we find a precise numerical agreement between the CAM calculation and the partial-wave calculation; this would not be the case if poles in this domain had been missed.

>>>>

Next, we examine the role of Regge poles in the time-independent scattering of monochromatic planar waves. We apply Complex Angular Momentum techniques to re-sum the partial wave series for the scattering amplitude, expressing it as a residue series evaluated at poles in the first quadrant, accompanied by a background integral. We compute the scattering cross section at several frequencies, and show precise agreement with the partial-wave calculations. Finally, we show that compact bodies naturally give rise to orbiting, glory, and rainbow-scattering interference effects.
\end{abstract}

@@

 \begin{figure}[htb]
\centering
 \includegraphics[scale=0.50]{potentials_s02_CD_R6L2_v3} \\ \vspace{0.2cm}
 \includegraphics[scale=0.50]{potentials_s02_CD_R2p26L2_v3}
\caption{\label{RP_approx_2Mw_3_6_s_1} The effective potential $V_\ell$ for a quadrupole ($\ell =2$) perturbation of a compact body of constant density and tenuity $R/M = 6$ (upper) and $R /M = 2.26$ (lower). The scalar field potential - Eq.~(\ref{Inside_Potentiel}) - and axial gravitational-wave potential - Eq.~(\ref{RW_Potentiel_axialGW}) - are indicated as solid/black lines and dotted/red lines, respectively. The horizontal axis is the tortoise coordinate $r_\ast/(2M)$ defined in Eq.~(\ref{eq:tortoise}).}
\label{fig:Veff}
\end{figure}

@@

Figure \ref{fig:rp1} shows the Regge pole spectrum for a neutron-star-like body ($R/M = 6$). We see that there are two branches of Regge poles in the first quadrant, which meet at $\text{Re}( \lambda_n(\omega) ) \sim \omega b_0$, where $b_0$ is the impact parameter for the null ray which grazes the surface of the compact body. The axial GW $s=2$ modes (filled markers) are typically close to their scalar-field $s=0$ counterparts (unfilled markers), as might be anticipated from the similarities in the effective potentials for the two cases (see Fig.~\ref{fig:Veff}).

\tom{I have added two comments here about the rest of the complex $\lambda$ plane. Feel free to edit.}

We find no evidence of Regge poles in the fourth quadrant when $\omega>0$. In addition, we find that $\lambda_n(-\omega) = \lambda_n(\omega)^*$ so that there are Regge poles in the fourth quadrant when $\omega<0$. Here we will only consider $\omega>0$.

We propose that the good analytical extension of modes into the second and third quadrants is to enforce the regular boundary condition for Re$[\lambda]<0$, i.e. $\phi_{\omega\ell} \sim r^{-\lambda+1/2}$ as $r\rightarrow 0$. Since the wave equation is invariant under the transformation $\lambda \rightarrow -\lambda$, it can be seen that the distribution of poles in the left half complex plane - Re$[\lambda]<0$ - follows from the poles in the right half complex plane by reflection through the origin.

@@

Now, we ``deform'' the contour ${\cal C}$ in Eq.~(\ref{SW_Scalar_Scattering_amp}) in order to collect, by using Cauchy's theorem, the Regge poles contributions. This is achieved by following, \textit{mutatis mutandis}, the approach developed in Ref~\cite{Folacci:2019cmc} (see more particularly Sec. IIB~3 and Fig.~1). We obtain
%
\begin{equation}\label{CAM_Scalar_Scattering_amp_tot}
\hat{f} (\omega, \theta) =  \hat{f}^\text{\tiny{B}} (\omega, \theta) +  \hat{f}^\text{\tiny{RP}} (\omega, \theta)
\end{equation}
where
\begin{subequations}\label{CAM_Scalar_Scattering_amp_decomp}
\begin{equation}\label{CAM_Scalar_Scattering_amp_decomp_Background}
\hat{f}^\text{\tiny{B}} (\omega, \theta) = \hat{f}^\text{\tiny{B},\tiny{Re}} (\omega, \theta)+ \hat{f}^\text{\tiny{B},\tiny{Im}} (\omega, \theta)
\end{equation}
is a background integral contribution with
\begin{equation}\label{CAM_Scalar_Scattering_amp_decomp_Background_a}
\hat{f}^\text{\tiny{B},\tiny{Re}} (\omega, \theta) = \frac{1}{\pi \omega} \int_{{\cal C}_{-}} d\lambda \, \lambda S_{\lambda -1/2}(\omega) Q_{\lambda -1/2}(\cos \theta +i0)
\end{equation}
and
\begin{eqnarray}\label{CAM_Scalar_Scattering_amp_decomp_Background_b}
\hat{f}^\text{\tiny{B},\tiny{Im}} && (\omega, \theta) = \frac{1}{2 \omega}\left(\int_{+i\infty}^{0} d\lambda \, \left[S_{\lambda -1/2}(\omega) P_{\lambda-1/2} (-\cos \theta) \right. \right.\nonumber \\
&& -\left. \left. S_{-\lambda -1/2}(\omega) e^{i \pi \left(\lambda+1/2\right)}P_{\lambda-1/2} (\cos \theta) \right]\lambda\right).
\end{eqnarray}
\end{subequations}
The second term in Eq.~(\ref{CAM_Scalar_Scattering_amp_tot}) ,
\begin{eqnarray}\label{CAM_Scalar_Scattering_amp_decomp_RP}
& & \hat{f}^\text{\tiny{RP}} (\omega, \theta) = -\frac{i \pi}{\omega}    \sum_{n=1}^{+\infty}   \frac{ \lambda_n(\omega) r_n(\omega)}{\cos[\pi \lambda_n(\omega)]}  \nonumber \\
&&  \qquad\qquad \qquad\qquad \times  P_{\lambda_n(\omega) -1/2} (-\cos \theta),
\end{eqnarray}
is a sum over the Regge poles lying in the first quadrant of the CAM plane. Of course, Eqs.~(\ref{CAM_Scalar_Scattering_amp_tot}), (\ref{CAM_Scalar_Scattering_amp_decomp}) and (\ref{CAM_Scalar_Scattering_amp_decomp_RP}) provide an exact representation of the scattering amplitude $\hat{f} (\omega, \theta)$ for the scalar field, equivalent to the initial partial wave expansion (\ref{Scalar_Scattering_amp}). From this CAM representation, we can extract the contribution $\hat{f}^\text{\tiny{RP}} (\omega, \theta)$ given by (\ref{CAM_Scalar_Scattering_amp_decomp_RP}) which, as a sum over Regge poles, is only an approximation of $\hat{f} (\omega, \theta)$, and which provides us with a corresponding approximation of the differential scattering cross section via (\ref{Scalar_Scattering_diff}).

@@

We can be confident that there are no additional branches of Regge poles in the first quadrant, for two reasons. First, we have scanned across the complex-$\lambda$ domains shown in Figs.~\ref{RP_R6_approx_2Mw_3_6_s_1} and \ref{RP_R226_approx_2Mw_3_6_s_1}, as detailed in Sec.~\ref{subsec:method}. Second, in our calculation of $d\sigma/d\Omega$, we find a precise numerical agreement between the CAM calculation and the partial-wave calculation; this would not be the case if poles in this domain had been missed.
